Active damping of rotating platforms using integral force feedback

### 2021-02-19 journal/paper.tex
Final submission

% Created 2020-11-02 lun. 15:33
% Intended LaTeX compiler: pdflatex
\documentclass[10pt]{iopart}

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{capt-of, subcaption}
\usepackage{hyperref}
\usepackage{amsmath, amssymb, amsfonts, cases, bm}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import}
\usepackage[USenglish, english]{babel}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{shapes.misc,arrows,arrows.meta}
\hypersetup{
 pdfauthor={Thomas Dehaeze},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.1.50 (Org mode 9.4)}, 
 pdflang={English}}
\begin{document}

\title{Active damping of rotating platforms using integral force feedback}
\author{Thomas Dehaeze$^{1,3}$ and Christophe Collette$^{1,2}$}
\address{$^1$ Precision Mechatronics Laboratory, University of Liege, Belgium}
\address{$^2$ BEAMS Department, Free University of Brussels, Belgium}
\address{$^3$ European Synchrotron Radiation Facility, Grenoble, France}
\ead{dehaeze.thomas@gmail.com}

\begin{abstract}
This paper investigates the use of Integral Force Feedback (IFF) for the active damping of rotating mechanical systems.
Guaranteed stability, typical benefit of IFF, is lost as soon as the system is rotating due to gyroscopic effects.
To overcome this issue, two modifications of the classical IFF control scheme are proposed.
The first consists of slightly modifying the control law while the second consists of adding springs in parallel with the force sensors.
Conditions for stability and optimal parameters are derived.
The results reveal that, despite their different implementations, both modified IFF control scheme have almost identical damping authority on suspension modes.
\end{abstract}

@@

\submitto{\SMS}
\maketitle
% \ioptwocol

\section{Introduction}
\label{sec:orgdf2365d}
\label{sec:introduction}
There is an increasing need to reduce the undesirable vibration of many sensitive equipment.
A common method to reduce vibration is to mount the sensitive equipment on a suspended platform which attenuates the vibrations above the frequency of the suspension modes.
In order to further decrease the residual vibrations, active damping can be used for reducing the magnification of the response in the vicinity of the resonances.

In \cite{preumont92_activ_dampin_by_local_force}, the Integral Force Feedback (IFF) control scheme has been proposed, where a force sensor, a force actuator and an integral controller are used to directly augment the damping of a mechanical system.
When the force sensor is collocated with the actuator, the open-loop transfer function has alternating poles and zeros which facilitate to guarantee the stability of the closed loop system \cite{preumont02_force_feedb_versus_accel_feedb}.

However, when the platform is rotating, gyroscopic effects alter the system dynamics and IFF cannot be applied as is.
The purpose of this paper is to study how the IFF strategy can be adapted to deal with these Gyroscopic effects.

The paper is structured as follows.
Section \ref{sec:dynamics} presents a simple model of a rotating suspended platform that will be used throughout this study.
Section \ref{sec:iff} explains how the unconditional stability of IFF is lost due to Gyroscopic effects induced by the rotation.
Section \ref{sec:iff_hpf} suggests a simple modification of the control law such that damping can be added to the suspension modes in a robust way.
Section \ref{sec:iff_kp} proposes to add springs in parallel with the force sensors to regain the unconditional stability of IFF.
Section \ref{sec:comparison} compares both proposed modifications to the classical IFF in terms of damping authority and closed-loop system behavior.

\section{Dynamics of Rotating Platforms}
\label{sec:orgb49fb34}
\label{sec:dynamics}
In order to study how the rotation does affect the use of IFF, a model of a suspended platform on top of a rotating stage is used.
Figure \ref{fig:system} represents the model schematically which is the simplest in which gyroscopic forces can be studied.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.9]{figs/fig01.pdf}
\caption{\label{fig:system}Schematic of the studied System}
\end{figure}

The rotating stage is supposed to be ideal, meaning it induces a perfect rotation \(\theta(t) = \Omega t\) where \(\Omega\) is the rotational speed in \(\si{\radian\per\second}\).

The suspended platform consists of two orthogonal actuators represented by three elements in parallel: a spring with a stiffness \(k\) in \(\si{\newton\per\meter}\), a dashpot with a damping coefficient \(c\) in \(\si{\newton\per\meter\second}\) and an ideal force source \(F_u, F_v\).
A payload with a mass \(m\) in \(\si{\kilo\gram}\), representing the sensitive equipment, is mounted on the (rotating) suspended platform.

@@

\par
To obtain the equations of motion for the system represented in Figure \ref{fig:system}, the Lagrangian equations are used:
\begin{equation}
\label{eq:lagrangian_equations}
  \frac{d}{dt} \left( \frac{\partial L}{\partial \dot{q}_i} \right) + \frac{\partial D}{\partial \dot{q}_i} - \frac{\partial L}{\partial q_i} = Q_i
\end{equation}
with \(L = T - V\) the Lagrangian, \(T\) the kinetic coenergy, \(V\) the potential energy, \(D\) the dissipation function, and \(Q_i\) the generalized force associated with the generalized variable \(\begin{bmatrix}q_1 & q_2\end{bmatrix} = \begin{bmatrix}d_u & d_v\end{bmatrix}\).
The equation of motion corresponding to the constant rotation in the \((\vec{i}_x, \vec{i}_y)\) plane is disregarded as the motion is considered to be imposed by the rotation stage.
\begin{equation}
\label{eq:energy_functions_lagrange}
  \begin{aligned}
    T &= \frac{1}{2} m \left( ( \dot{d}_u - \Omega d_v )^2 + ( \dot{d}_v + \Omega d_u )^2 \right), \\
    V &= \frac{1}{2} k \big( {d_u}^2 + {d_v}^2 \big), \ Q_1 = F_u, \\
    D &= \frac{1}{2} c \big( \dot{d}_u{}^2 + \dot{d}_v{}^2 \big), \ Q_2 = F_v
  \end{aligned}
\end{equation}

Substituting \eqref{eq:energy_functions_lagrange} into \eqref{eq:lagrangian_equations} for both generalized coordinates gives two coupled differential equations
\begin{subequations}
\label{eq:eom_coupled}
  \begin{align}
    m \ddot{d}_u + c \dot{d}_u + ( k - m \Omega^2 ) d_u &= F_u + 2 m \Omega \dot{d}_v \\
    m \ddot{d}_v + c \dot{d}_v + ( k \underbrace{-\,m \Omega^2}_{\text{Centrif.}} ) d_v &= F_v \underbrace{-\,2 m \Omega \dot{d}_u}_{\text{Coriolis}}
  \end{align}
\end{subequations}

The uniform rotation of the system induces two Gyroscopic effects as shown in \eqref{eq:eom_coupled}:
\begin{itemize}
\item Centrifugal forces: that can been seen as added negative stiffness \(- m \Omega^2\) along \(\vec{i}_u\) and \(\vec{i}_v\)
\item Coriolis Forces: that couples the motion in the two orthogonal directions
\end{itemize}

One can verify that without rotation (\(\Omega = 0\)) the system becomes equivalent to two uncoupled one degree of freedom mass-spring-damper systems:
\begin{subequations}
\label{eq:oem_no_rotation}
  \begin{align}
    m \ddot{d}_u + c \dot{d}_u + k d_u &= F_u \\
    m \ddot{d}_v + c \dot{d}_v + k d_v &= F_v
  \end{align}
\end{subequations}

\par
To study the dynamics of the system, the differential equations of motions \eqref{eq:eom_coupled} are transformed in the Laplace domain and the \(2 \times 2\) transfer function matrix \(\mathbf{G}_d\) from \(\begin{bmatrix}F_u & F_v\end{bmatrix}\) to \(\begin{bmatrix}d_u & d_v\end{bmatrix}\) is obtained.

@@

\begin{subequations}
\begin{align}
  \mathbf{G}_{d}(1,1) &= {\frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}} \\
                  &= \mathbf{G}_{d}(2,2) \nonumber \\
  \mathbf{G}_{d}(1,2) &= {\frac{2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}} \\
                  &= -\mathbf{G}_{d}(1,2) \nonumber
\end{align}
\end{subequations}

To simplify the analysis, the undamped natural frequency \(\omega_0\) and the damping ratio \(\xi\) are used
\begin{equation}
  \omega_0 = \sqrt{\frac{k}{m}} \text{ in } \si{\radian\per\second}, \quad \xi = \frac{c}{2 \sqrt{k m}}
\end{equation}

The transfer function matrix \(\mathbf{G}_d\) becomes equal to
\begin{subequations}
\label{eq:Gd_w0_xi_k}
  \begin{align}
    \mathbf{G}_{d}(1,1) &= {\scriptstyle \frac{\frac{1}{k} \left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} } \\
    \mathbf{G}_{d}(1,2) &= {\scriptstyle \frac{\frac{1}{k} \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} }
  \end{align}
\end{subequations}

@@

\par
The poles of \(\mathbf{G}_d\) are the complex solutions \(p\) of \eqref{eq:poles}.

@@

Supposing small damping (\(\xi \ll 1\)), two pairs of complex conjugate poles are obtained \eqref{eq:pole_values}.

@@

The real part and complex part of these two pairs of complex conjugate poles are represented in Figure \ref{fig:campbell_diagram} as a function of the rotational speed \(\Omega\).
As the rotational speed increases, \(p_{+}\) goes to higher frequencies and \(p_{-}\) to lower frequencies.
The system becomes unstable for \(\Omega > \omega_0\) as the real part of \(p_{-}\) is positive.
Physically, the negative stiffness term \(-m\Omega^2\) induced by centrifugal forces exceeds the spring stiffness \(k\).

@@

\begin{figure}[htbp]
\begin{subfigure}[c]{0.48\linewidth}
\includegraphics[width=\linewidth]{figs/fig02a.pdf}
\caption{\label{fig:campbell_diagram_real} Real Part}
\end{subfigure}
\hfill
\begin{subfigure}[c]{0.48\linewidth}
\includegraphics[width=\linewidth]{figs/fig02b.pdf}
\caption{\label{fig:campbell_diagram_imag} Imaginary Part}
\end{subfigure}
\hfill
\caption{\label{fig:campbell_diagram}Campbell Diagram : Evolution of the complex and real parts of the system's poles as a function of the rotational speed \(\Omega\)}
\centering
\end{figure}

Looking at the transfer function matrix \(\mathbf{G}_d\) in \eqref{eq:Gd_w0_xi_k}, one can see that the two diagonal (direct) terms are equal and the two off-diagonal (coupling) terms are opposite.
The bode plot of these two terms are shown in Figure \ref{fig:plant_compare_rotating_speed} for several rotational speeds \(\Omega\).
These plots confirm the expected behavior: the frequency of the two pairs of complex conjugate poles are further separated as \(\Omega\) increases.
For \(\Omega > \omega_0\), the low frequency pair of complex conjugate poles \(p_{-}\) becomes unstable.

\begin{figure}[htbp]
\begin{subfigure}[c]{0.48\linewidth}
\includegraphics[width=\linewidth]{figs/fig03a.pdf}
\caption{\label{fig:plant_compare_rotating_speed_direct} Direct Terms}
\end{subfigure}
\hfill
\begin{subfigure}[c]{0.48\linewidth}
\includegraphics[width=\linewidth]{figs/fig03b.pdf}
\caption{\label{fig:plant_compare_rotating_speed_coupling} Coupling Terms}
\end{subfigure}
\hfill
\caption{\label{fig:plant_compare_rotating_speed}Bode Plots for \(\mathbf{G}_d\) for several rotational speed \(\Omega\)}
\centering
\end{figure}

\section{Decentralized Integral Force Feedback}
\label{sec:org7b1adf9}
\label{sec:iff}
In order to apply IFF to the system, force sensors are added in series with the two actuators (Figure \ref{fig:system_iff}).
As this study focuses on decentralized control, two identical controllers \(K_F\) are used to feedback each of the sensed force to its associated actuator and no attempt is made to counteract the interactions in the system.
The control diagram is schematically shown in Figure \ref{fig:control_diagram_iff}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.9]{figs/fig04.pdf}
\caption{\label{fig:system_iff}System with added Force Sensor in series with the actuators}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/fig05.pdf}
\caption{\label{fig:control_diagram_iff}Control Diagram for decentralized IFF}
\end{figure}

\par
The forces \(\begin{bmatrix}f_u & f_v\end{bmatrix}\) measured by the two force sensors represented in Figure \ref{fig:system_iff} are equal to
\begin{equation}
\label{eq:measured_force}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} =
  \begin{bmatrix} F_u \\ F_v \end{bmatrix} - (c s + k)
  \begin{bmatrix} d_u \\ d_v \end{bmatrix}
\end{equation}

Inserting \eqref{eq:Gd_w0_xi_k} into \eqref{eq:measured_force} yields
\begin{equation}
\label{eq:Gf_mimo_tf}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} = \mathbf{G}_{f} \begin{bmatrix} F_u \\ F_v \end{bmatrix}
\end{equation}

\begin{subequations}
\label{eq:Gf}
  \begin{align}
    \mathbf{G}_{f}(1,1) &= {\scriptstyle \frac{\left( \frac{s^2}{{\omega_0}^2} - \frac{\Omega^2}{{\omega_0}^2} \right) \left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right) + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} } \\
    \mathbf{G}_{f}(1,2) &= {\scriptstyle \frac{- \left( 2 \xi \frac{s}{\omega_0} + 1 \right) \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} }
  \end{align}
\end{subequations}

The zeros of the diagonal terms of \(\mathbf{G}_f\) are equal to (neglecting the damping for simplicity)
\begin{subequations}
  \begin{align}
    z_c &= \pm j \omega_0 \sqrt{\frac{1}{2} \sqrt{8 \frac{\Omega^2}{{\omega_0}^2} + 1} + \frac{\Omega^2}{{\omega_0}^2} + \frac{1}{2} } \label{eq:iff_zero_cc} \\
    z_r &= \pm   \omega_0 \sqrt{\frac{1}{2} \sqrt{8 \frac{\Omega^2}{{\omega_0}^2} + 1} - \frac{\Omega^2}{{\omega_0}^2} - \frac{1}{2} } \label{eq:iff_zero_real}
  \end{align}
\end{subequations}

>>>>

% Created 2021-02-19 ven. 11:49
% Intended LaTeX compiler: pdflatex
\documentclass[10pt]{iopart}

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{capt-of, subcaption}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{amsmath, amssymb, amsfonts, cases, bm}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import}
\usepackage[USenglish, english]{babel}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{shapes.misc,arrows,arrows.meta}
\usepackage[square,comma,sort&compress,numbers]{natbib}
\renewcommand{\bibsection}{\section*{References}}
\hypersetup{
 pdfauthor={Thomas Dehaeze},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.1 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\title{Active damping of rotating platforms using integral force feedback}
\author{Thomas Dehaeze$^{1,3}$ and Christophe Collette$^{1,2}$}
\address{$^1$ Precision Mechatronics Laboratory, University of Liege, Belgium}
\address{$^2$ BEAMS Department, Free University of Brussels, Belgium}
\address{$^3$ European Synchrotron Radiation Facility, Grenoble, France}
\ead{tdehaeze@uliege.be}

\begin{abstract}
This paper investigates the use of Integral Force Feedback (IFF) for the active damping of rotating mechanical systems.
Guaranteed stability, typical benefit of IFF, is lost as soon as the system is rotating due to gyroscopic effects.
To overcome this issue, two modifications of the classical IFF control scheme are proposed.
The first consists of slightly modifying the control law while the second consists of adding springs in parallel with the force sensors.
Conditions for stability and optimal parameters are derived.
The results reveal that, despite their different implementations, both modified IFF control scheme have almost identical damping authority on the suspension modes.
\end{abstract}

@@

\submitto{{\it Engineering Research Express\/}}
\maketitle
% Two Columns output
\ioptwocol

\section{Introduction}
\label{sec:org1def047}
\label{sec:introduction}
There is an increasing need to reduce the undesirable vibration of sensitive equipment in many different fields such as in aerospace industry~\cite{hanieh03_activ_stewar,hauge04_sensor_contr_space_based_six,souleille18_concep_activ_mount_space_applic}, semi conductor industry~\cite{yoshioka01_activ_microv_isolat_system_hi}, microscopy~\cite{lan08_activ_vibrat_isolat_long_range,fleming15_low_order_dampin_track_contr}, gravitational wave detectors~\cite{matichard15_seism_isolat_advan_ligo} and particles accelerators~\cite{collette10_activ_quadr_stabil_futur_linear_partic_collid,dehaeze18_sampl_stabil_for_tomog_exper}.

A common method to reduce vibration is to mount the sensitive equipment on a suspended platform which attenuates the vibrations above the frequency of the suspension modes~\cite{reilly06_critic,poel10_explor_activ_hard_mount_vibrat}. \par
In order to further decrease the residual vibrations, active damping can be used for reducing the magnification of the response in the vicinity of the resonances~\cite{collette11_review_activ_vibrat_isolat_strat}.

Many active damping techniques have been developed over the years such as Positive Position Feedback (PPF)~\cite{lin06_distur_atten_precis_hexap_point,fanson90_posit_posit_feedb_contr_large_space_struc}, Integral Force Feedback (IFF)~\cite{preumont91_activ} and Direct Velocity Feedback (DVF)~\cite{karnopp74_vibrat_contr_using_semi_activ_force_gener,serrand00_multic_feedb_contr_isolat_base_excit_vibrat,preumont02_force_feedb_versus_accel_feedb}. \par
In~\cite{preumont92_activ_dampin_by_local_force}, the IFF control scheme has been proposed, where a force sensor, a force actuator and an integral controller are used to directly augment the damping of a mechanical system.
When the force sensor is collocated with the actuator, the open-loop transfer function has alternating poles and zeros which facilitate to guarantee the stability of the closed loop system~\cite{preumont02_force_feedb_versus_accel_feedb}.
It was latter shown that this property holds for multiple collated actuator/sensor pairs~\cite{preumont08_trans_zeros_struc_contr_with}. \par
The main advantages of IFF over other active damping techniques are the guaranteed stability even in presence of flexible dynamics, good performances and robustness properties~\cite{preumont02_force_feedb_versus_accel_feedb}. \par
Several improvements of the classical IFF have been proposed, such as adding a feed-through term to increase the achievable damping~\cite{teo15_optim_integ_force_feedb_activ_vibrat_contr} or adding an high pass filter to recover the loss of compliance at low frequency~\cite{chesne16_enhan_dampin_flexib_struc_using_force_feedb}.
Recently, an \(\mathcal{H}_\infty\) optimization criterion has been used to derive optimal gains for the IFF controller~\cite{zhao19_optim_integ_force_feedb_contr}. \par
However, when the platform is rotating, gyroscopic effects alter the system dynamics and IFF cannot be applied as is.
The purpose of this paper is to study how the IFF strategy can be adapted to deal with these gyroscopic effects. \par
The paper is structured as follows.
Section~\ref{sec:dynamics} presents a simple model of a rotating suspended platform that will be used throughout this study.
Section~\ref{sec:iff} explains how the unconditional stability of IFF is lost due to gyroscopic effects induced by the rotation.
Section~\ref{sec:iff_hpf} suggests a simple modification of the control law such that damping can be added to the suspension modes in a robust way.
Section~\ref{sec:iff_kp} proposes to add springs in parallel with the force sensors to regain the unconditional stability of IFF.
Section~\ref{sec:comparison} compares both proposed modifications to the classical IFF in terms of damping authority and closed-loop system behavior.

\section{Dynamics of Rotating Platforms}
\label{sec:org414e8cb}
\label{sec:dynamics}
In order to study how the rotation affects the use of IFF, a model of a suspended platform on top of a rotating stage is used.
Figure~\ref{fig:system} represents a schematic of the model which is the simplest in which gyroscopic forces can be studied.

\begin{figure}[htbp]
\centering
\includegraphics[scale=1,scale=0.9]{figs/fig01.pdf}
\caption{\label{fig:system}Schematic of the studied system}
\end{figure}

The rotating stage is supposed to be ideal, meaning it induces a perfect rotation \(\theta(t) = \Omega t\) where \(\Omega\) is the rotational speed in \(\si{\radian\per\s}\).

The suspended platform consists of two orthogonal actuators each represented by three elements in parallel: a spring with a stiffness \(k\) in \(\si{\newton\per\meter}\), a dashpot with a damping coefficient \(c\) in \(\si{\newton\per(\meter\per\second)^{-1}}\) and an ideal force source \(F_u, F_v\).
A payload with a mass \(m\) in \(\si{\kilo\gram}\), representing the sensitive equipment, is mounted on the (rotating) suspended platform.

@@

\par
To obtain the equations of motion for the system represented in Figure~\ref{fig:system}, the Lagrangian equations are used:
\begin{equation}
\label{eq:lagrangian_equations}
  \frac{d}{dt} \left( \frac{\partial L}{\partial \dot{q}_i} \right) + \frac{\partial D}{\partial \dot{q}_i} - \frac{\partial L}{\partial q_i} = Q_i
\end{equation}
with \(L = T - V\) the Lagrangian, \(T\) the kinetic coenergy, \(V\) the potential energy, \(D\) the dissipation function, and \(Q_i\) the generalized force associated with the generalized variable \(\begin{bmatrix}q_1 & q_2\end{bmatrix} = \begin{bmatrix}d_u & d_v\end{bmatrix}\).
The equation of motion corresponding to the constant rotation along \(\vec{i}_w\) is disregarded as this motion is considered to be imposed by the rotation stage.
\begin{equation}
\label{eq:energy_functions_lagrange}
  \begin{aligned}
    T &= \frac{1}{2} m \left( ( \dot{d}_u - \Omega d_v )^2 + ( \dot{d}_v + \Omega d_u )^2 \right), \\
    V &= \frac{1}{2} k \big( {d_u}^2 + {d_v}^2 \big), \ Q_1 = F_u, \\
    D &= \frac{1}{2} c \big( \dot{d}_u{}^2 + \dot{d}_v{}^2 \big), \ Q_2 = F_v
  \end{aligned}
\end{equation}

Substituting Eq.~\eqref{eq:energy_functions_lagrange} into Eq.~\eqref{eq:lagrangian_equations} for both generalized coordinates gives two coupled differential equations~\eqref{eq:eom_coupled_1} and~\eqref{eq:eom_coupled_2}.
\begin{subequations}
\label{eq:eom_coupled}
  \begin{align}
    m \ddot{d}_u + c \dot{d}_u + ( k - m \Omega^2 ) d_u &= F_u + 2 m \Omega \dot{d}_v \label{eq:eom_coupled_1} \\
    m \ddot{d}_v + c \dot{d}_v + ( k \underbrace{-\,m \Omega^2}_{\text{Centrif.}} ) d_v &= F_v \underbrace{-\,2 m \Omega \dot{d}_u}_{\text{Coriolis}} \label{eq:eom_coupled_2}
  \end{align}
\end{subequations}

The uniform rotation of the system induces two gyroscopic effects as shown in Eq.~\eqref{eq:eom_coupled}:
\begin{itemize}
\item Centrifugal forces: that can been seen as an added negative stiffness \(- m \Omega^2\) along \(\vec{i}_u\) and \(\vec{i}_v\)
\item Coriolis Forces: that adds coupling between the two orthogonal directions.
\end{itemize}

One can verify that without rotation (\(\Omega = 0\)) the system becomes equivalent to two uncoupled one degree of freedom mass-spring-damper systems.

\par
To study the dynamics of the system, the differential equations of motions~\eqref{eq:eom_coupled} are converted into the Laplace domain and the \(2 \times 2\) transfer function matrix \(\mathbf{G}_d\) from \(\begin{bmatrix}F_u & F_v\end{bmatrix}\) to \(\begin{bmatrix}d_u & d_v\end{bmatrix}\) in Eq.~\eqref{eq:Gd_mimo_tf} is obtained.
Its elements are shown in Eq.~\eqref{eq:Gd_indiv_el}.

@@

\begin{subequations}
\label{eq:Gd_indiv_el}
\begin{align}
  & \mathbf{G}_{d}(1,1) = \mathbf{G}_{d}(2,2) = \dots \nonumber \\
  & \quad {\frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}} \\
  & \mathbf{G}_{d}(1,2) = -\mathbf{G}_{d}(1,2) = \dots \nonumber \\
  & \quad {\frac{2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}}
\end{align}
\end{subequations}

To simplify the analysis, the undamped natural frequency \(\omega_0\) and the damping ratio \(\xi\) are used as in Eq.~\eqref{eq:xi_and_omega}.
\begin{equation}
\label{eq:xi_and_omega}
  \omega_0 = \sqrt{\frac{k}{m}} \text{ in } \si{\radian\per\second}, \quad \xi = \frac{c}{2 \sqrt{k m}}
\end{equation}

The elements of transfer function matrix \(\mathbf{G}_d\) are now describe by Eq.~\eqref{eq:Gd_w0_xi_k}.
\begin{subequations}
\label{eq:Gd_w0_xi_k}
  \begin{align}
    \mathbf{G}_{d}(1,1) &= {\scriptstyle \frac{\frac{1}{k} \left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2}}  \\
    \mathbf{G}_{d}(1,2) &= {\scriptstyle \frac{\frac{1}{k} \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2}}
  \end{align}
\end{subequations}

@@

\par
The poles of \(\mathbf{G}_d\) are the complex solutions \(p\) of Eq.~\eqref{eq:poles}.

@@

Supposing small damping (\(\xi \ll 1\)), two pairs of complex conjugate poles are obtained as shown in Eq.~\eqref{eq:pole_values}.

@@

The real and complex parts of these two pairs of complex conjugate poles are represented in Figure~\ref{fig:campbell_diagram} as a function of the rotational speed \(\Omega\).
As the rotational speed increases, \(p_{+}\) goes to higher frequencies and \(p_{-}\) goes to lower frequencies.
The system becomes unstable for \(\Omega > \omega_0\) as the real part of \(p_{-}\) is positive.
Physically, the negative stiffness term \(-m\Omega^2\) induced by centrifugal forces exceeds the spring stiffness \(k\).

@@

\begin{figure}[htbp]
\begin{subfigure}[c]{0.48\linewidth}
\centering
\includegraphics[scale=1]{figs/fig02a.pdf}
\caption{\label{fig:campbell_diagram_real} Real Part}
\end{subfigure}
\hfill
\begin{subfigure}[c]{0.48\linewidth}
\centering
\includegraphics[scale=1]{figs/fig02b.pdf}
\caption{\label{fig:campbell_diagram_imag} Imaginary Part}
\end{subfigure}
\hfill
\caption{\label{fig:campbell_diagram}Campbell diagram : Evolution of the complex and real parts of the system's poles as a function of the rotational speed \(\Omega\)}
\centering
\end{figure}

Looking at the transfer function matrix \(\mathbf{G}_d\) in Eq.~\eqref{eq:Gd_w0_xi_k}, one can see that the two diagonal (direct) terms are equal and that the two off-diagonal (coupling) terms are opposite.
The bode plot of these two terms are shown in Figure~\ref{fig:plant_compare_rotating_speed} for several rotational speeds \(\Omega\).
These plots confirm the expected behavior: the frequency of the two pairs of complex conjugate poles are further separated as \(\Omega\) increases.
For \(\Omega > \omega_0\), the low frequency pair of complex conjugate poles \(p_{-}\) becomes unstable.

\begin{figure}[htbp]
\begin{subfigure}[c]{0.48\linewidth}
\centering
\includegraphics[scale=1]{figs/fig03a.pdf}
\caption{\label{fig:plant_compare_rotating_speed_direct} Direct Terms}
\end{subfigure}
\hfill
\begin{subfigure}[c]{0.48\linewidth}
\centering
\includegraphics[scale=1]{figs/fig03b.pdf}
\caption{\label{fig:plant_compare_rotating_speed_coupling} Coupling Terms}
\end{subfigure}
\hfill
\caption{\label{fig:plant_compare_rotating_speed}Bode plots for \(\mathbf{G}_d\) for several rotational speed \(\Omega\)}
\centering
\end{figure}

\section{Decentralized Integral Force Feedback}
\label{sec:orgdc87c6d}
\label{sec:iff}
In order to apply IFF to the rotating system, force sensors are added in series with the two actuators (Figure~\ref{fig:system_iff}).
As this study focuses on decentralized control, two identical controllers \(K_F\) are used to feedback each of the sensed force to its associated actuator and no attempt is made to counteract the interactions in the system.
The control diagram is schematically shown in Figure~\ref{fig:control_diagram_iff}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=1,scale=0.9]{figs/fig04.pdf}
\caption{\label{fig:system_iff}System with added force sensor in series with the actuators and with a decentralized IFF architecture}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=1,scale=1]{figs/fig05.pdf}
\caption{\label{fig:control_diagram_iff}Control diagram for decentralized IFF}
\end{figure}

\par
The forces \(\begin{bmatrix}f_u & f_v\end{bmatrix}\) measured by the two force sensors represented in Figure~\ref{fig:system_iff} are described by Eq.~\eqref{eq:measured_force}.
\begin{equation}
\label{eq:measured_force}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} =
  \begin{bmatrix} F_u \\ F_v \end{bmatrix} - (c s + k)
  \begin{bmatrix} d_u \\ d_v \end{bmatrix}
\end{equation}

The transfer function matrix \(\mathbf{G}_{f}\) from actuator forces to measured forces in Eq.~\eqref{eq:Gf_mimo_tf} can be obtained by inserting Eq.~\eqref{eq:Gd_w0_xi_k} into Eq.~\eqref{eq:measured_force}.
Its elements are shown in Eq.~\eqref{eq:Gf_indiv_el}.

\begin{equation}
\label{eq:Gf_mimo_tf}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} = \mathbf{G}_{f} \begin{bmatrix} F_u \\ F_v \end{bmatrix}
\end{equation}

\begin{subequations}
\label{eq:Gf_indiv_el}
\label{eq:Gf}
  \begin{align}
    & \mathbf{G}_{f}(1,1) = \mathbf{G}_{f}(2,2) = \dots \nonumber \\
    & \quad {\scriptstyle \frac{\left( \frac{s^2}{{\omega_0}^2} - \frac{\Omega^2}{{\omega_0}^2} \right) \left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right) + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} } \label{eq:Gf_diag_tf} \\
    & \mathbf{G}_{f}(1,2) = -\mathbf{G}_{f}(2,1) = \dots \nonumber \\
    & \quad {\scriptstyle \frac{- \left( 2 \xi \frac{s}{\omega_0} + 1 \right) \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)}{\left( \frac{s^2}{{\omega_0}^2} + 2 \xi \frac{s}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{s}{\omega_0} \right)^2} } \label{eq:Gf_off_diag_tf}
  \end{align}
\end{subequations}
